---
title: "Midterm One"
author: "Aimee A"
date: "February 2, 2022"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

_In this midterm we will analyze some data on the conservation status of species in North America and spending under the Endangered Species Act._

_First read in the file conservationdata.csv_

```{r}
cons_dat=read.csv("/home/aimee/Documents/ESP_106/Midterm/conservationdata.csv")

```

_1. What fraction of species in the dataset are listed under the Endangered Species Act? (2 points)_

```{r}
esa1=sum(cons_dat$listed==1)
total=dim(cons_dat)
esa1/total[1]
```
**Answer:** 1,617/53,638 or about 3% species are listed as endangered.

_2. Make a table (i.e. data frame) giving the breakdown of species by taxonomic group (i.e. one column is the taxonomic group and the second column is what fraction of the dataset that group represents) (4 points)_

```{r}
tax=unique(cons_dat$taxon)
df_tax=data.frame(tax)
fract=tax[0]


for (i in tax[1:9]) {
  x=sum(cons_dat$tax==i)
   fract[[i]]=(sum(cons_dat$taxon==i)/total[1])
}

df_tax$rep=fract
print(df_tax)

```

_3. a) One interesting question is how the conservation status varies between different taxonomic groups. Make a plot showing the distribution of conservation status within each taxonomic group (3 points)_

```{r}
library(ggplot2)
library(RColorBrewer)
ggplot(cons_dat, aes(x=taxon, fill=conservation_status))+
  geom_bar(stat = "count")+
ggtitle("Conservation Status by Taxonomic Group")+
  xlab("Taxonomic Group") + ylab("Count")+
  theme(axis.text.x = element_text(angle = 35), plot.title = element_text(hjust = 0.5))+
  scale_fill_brewer(palette = "Dark2",name="Conservation Status",
                    labels=c("Critically Imperiled","Imperiled","Vulnerable","Apparently Secure","Secure","Extinct","Probably Extinct","Unkown","Not Listed"))

```


_b) Based on this graph, what is something we might be concerned about in terms of analyzing the data on conservation status, particularly for fungi and invertebrates? (1 point)_

**Answer:** Both these groups have a high number of unknown conservation statuses. 

_Now read in the second data file, spendingdata.csv_
```{r}
spnd=read.csv("/home/aimee/Documents/ESP_106/Midterm/spendingdata.csv")
```

_4. Merge in the data from the conservation status data frame to the spending data frame, so that we have information on species names, taxonomic group, and conservation status with the spending data. (2 points)_

```{r}
df_m=merge(cons_dat,spnd, by="speciesid")
```

_5. a) Make a plot showing the distribution of spending in the year 2016 (3 points)_

```{r}
y16=df_m[(df_m$Year=="2016"),]

ggplot(y16,aes(y=spending),col=speciesname)+
  geom_density(stat = "bin",color="darkblue")+
  xlab("Species Count") + ylab(label = NULL)+
  scale_y_continuous(n.breaks = 5,labels = scales::dollar_format())+
   coord_flip()+
  ggtitle("Spending Distribution for 2016")+
  theme(
    plot.title = element_text(hjust = 0.5, size = 12)
  )
  
```

_b) Notice the (very) long right tail on spending data - we spend a lot on a very small number of species. Identify the 3 species with the most spending in 2016. (2 points)_

```{r}
y16sort=y16[order(y16$spending,decreasing = TRUE),]
y16sort=y16sort[,-c(1)]
#I removed species id to make the table fit nicely onto the html page. 
print(y16sort[c(1:3),])
```
_Look up these scientific names - what type of species are these?_

**Answer:** Chinook Salmon, Rainbow Trout, Coho Salmon 

_6. Finally, we will use a regression to look at the relationship between spending and species taxon._

_a) Drop the rows where spending == 0 from the data frame and then make a new column with the logarithm (log()) of spending in each year. (2 points)_

```{r}
omit0=which(df_m$spending==0)
yall=df_m[-c(omit0),]
yall$based=log(yall$spending)
yy16=yall[(yall$Year=="2016"),]

ggplot(yy16,aes(y=based),col=speciesname)+
  geom_density(stat = "bin",color="darkblue")+
  xlab("Species Count") + ylab(label = "Spending (Log)")+
  ggtitle("Spending Distribution for 2016", subtitle= "*Adjusted")+
    coord_flip()+
  theme(
    plot.title = element_text(hjust = 0.5, size = 12),
    plot.subtitle = element_text(hjust = .5,size=10)
    )
```


_b) Run a regression of logged spending on taxonomic group and print the summary for the regression below (3 points)_

```{r}
lreg=lm(based~taxon, yall)
summary(lreg)
```
_c) Based on your results in b, do we see statistically significant differences in spending between different taxonomic groups? If so, which kinds of species tend to have more spent on them and which have less? (1 points_

**Answer:** Difference in spending between different taxonomic groups is statistically significant, (p-value < 0.5). Mammals tend to receive more funding, plants tend to receive the least amount of funding. 

_7. Push your R markdown file to your Github repository (2 points)_
