---
title: "BLM_USGS"
author: "mmz"
date: "4/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir="/home/aimee/Documents/BLM_USGS/Coding_Analysis/")
library(readxl)
library(stringr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggcorrplot)
```

## Preliminary Coding Analysis for CEA Project

Main Objectives:
1. Basic Number Counts
  - barplot for resource counts
  - average resouces with CE per EA
2. Word Clouds for Language Analysis
  - question 21 and notes
3. Question Counts / Tier Analysis  
  - box plot for highest frequency yes questions
  - yes counts per question with respect to resource and also region
  -->Correlleogram 
4. Statistical Significance
  - running model to determine if one question in particular is an indicator of
  highly rated eas (>3stars)
  - running methods analysis, which questions were most impactful 

5/2 To Do List:
1. Incorporate prelim data cleaning chunk and run through
2. Prelim charts, annotate and run through. Should include the following:
  a- Table distribution of sample (Jen's stratified sample)
  b- **If time allows: map with sampled states/regions and # sampled there
  Resource Graph, count of resources analyzed
  c- Resource Count w/ percent having CE analyzed (bar plot, ordered by highest CE)
  d- Maybe finalize resource by yes/no/NAs, not sure how helpful this is
  e- Identify top teired EAs, these will have highest star rating ***DO THIS FIRST, ANALYSIS TEAM CAN PULL AND CATHER DISCUSSION POINTS FOR THESE
3. CE breakdowns, look only at those EAs with analyzed CEs and explore the following:
  a- yes area line graph
  b- most conclusively no questions, questions ordered by most no responses
  c- identify outliers for closer examination
  **DO OUTLIERS FIRST FOR ANALYSIS BY EXEC TEAM
4. Broad analysis breakdown and statistical question analysis
  a- take average of yelp reviews and compare with broad rating/populate broads with na rating
  b- analyze broad by resource
  c- try to find "choke point questions" or gate keeper questions with strong implications for overall EA quality
  d- lumped, present/absent/categorized analysis 
  
### TWO MAIN OBJECTIVES:
- highlight clear and consistent work being done by BLM staff within the cumulative effects section of EA's
- identifying existing good examples of cumulatives effects 


  
# In this chunk we will focus on cleaning and standardizing the data. 

```{r}

# Read in  excel file.
datsheet=readxl::read_excel("Master_Sheet.xlsx")

# Rename columns for convenience. 
colnames(datsheet)=c("Coder","EA_Number","EA_Name","Resource_Given","Resource_General","Broad_1","Broad_2","Broad_3", "Resource_4", 5:21, "21B",22, "22B", 23:26, "Notes","Stars","Project_Type")

# Remove empty columns/rows.
datsheet=datsheet[-c(1),-c(37)]


# Creating a state column. Create a combined column to standardize resources. 

## Though we listed some unique resources in the Resource_given column, we categorized them according to what standard resource they would likely fall under in the Resource_general column (e.g, unique Resource_given: "User Experience", Resource_general: Recreation). 

datsheet=datsheet%>%
  mutate(ST=substr(EA_Number,9,10))%>%
  mutate(Resource_Main=coalesce(Resource_General,Resource_Given))

# Standardizing NA values.
datsheet[is.na(datsheet)]= "N/A"

# Standardizing capitalization to all caps. 
datsheet= mutate_all(datsheet,toupper)

#Data is clean. Datsheet will remain base data frame with no further edits. 
```

```{r, resource count}
# We are filtering 

nobroad=datsheet%>%
  filter(!Resource_Main=="BROAD")%>%
  group_by(Resource_Main)%>%
  mutate(Rcount=n())


resource_count=ggplot(nobroad,aes(x=reorder(Resource_Main,Rcount)))+
  geom_bar()+
  theme_classic()+
  coord_flip()

resource_count
```


```{r, response count area graph} 

#yes/no/na count by resource type

count_No=function(nobroad) sum(nobroad[,(9:33)]=="NO")
count_Yes=function(nobroad) sum(nobroad[,(9:33)]=="YES")
count_NA=function(nobroad) sum(nobroad[,(9:33)]=="N/A")


response_count=plyr::ddply(nobroad,"Resource_Main",c(nope= count_No, Yes= count_Yes, notapp=count_NA))

#there is a great variation in responses for each resource. therefore, we will mutate the response df to indicate percentage of total questions rather than count.

response_count=response_count%>%
  mutate(rsum=rowSums(select(.,(2:4))))%>%
  mutate("Y"=Yes/rsum,"N"=nope/rsum,"NullValues"=notapp/rsum)

response_count=response_count%>%
  select(.,c(1,6:8))%>%
  pivot_longer(!Resource_Main,names_to = "Response",values_to = "percent")%>%
  group_by(Resource_Main)%>%
  mutate(Rcount=n())

View(response_count)

#plot to show composition of question response by resource

qper=ggplot(response_count, aes(x=Resource_Main, y=percent, fill= Response))+
  geom_bar(stat="identity")+
  scale_fill_manual("Response Type",values = c("Y"="darkgreen","N"="red","NullValues"="goldenrod1"))+
  coord_flip()

qper

```

```{r, resources by frequency of CE analyzed}

#counting frequency of cumulative effects analyzed

CEonly=sum(datsheet$`Resource_4`=="Yes")/length(datsheet$`Resource_4`)

```

```{r, broad category analysis}

broad=datsheet%>%
  filter(Resource_Main=="BROAD")

broad_No=function(broad) sum(broad[,(6:8)]=="NO")
broad_Yes=function(broad) sum(broad[,(6:8)]=="YES")

broad_count=plyr::ddply(broad,"Project_Type",c(nope= broad_No, Yes= broad_Yes))
broad_count=broad_count%>%
  pivot_longer(names_to = "response",values_to = nope,Yes)
  
View(broad_count)

broad_count=ggplot(broad_count, aes(x=Project_Type))+
  geom_bar(stat="identity")
broad_count
```


```{r, area line chart for each question response}

```

```{r, top tier analysis}
# For example sampling

toptier=datsheet%>%
  filter(!Stars=="N/A")%>%
  filter(Stars>="4â­")

save(toptier,file="toptier.rdata")
write.csv(toptier,"toptier.csv", row.names = FALSE)
```

```{r, CE only analysis}
CEonly=datsheet%>%
  filter(`Resource_4`== "YES")
```

```{r, Correlations}
#
#Begin Jim edits
#correlation using complete observations (no NAs) 
#helper function
helperFunction <- function(x){
    ifelse(x=="YES", 1,0)
}
#subset datasheet for columns 5 through 10 and change YES to 1 everything else to 0
df <- datsheet%>% 
  select('5','6','7','8','9', '10','22')%>%
  filter(`5`!= "N/A")%>%
  mutate_all(helperFunction)
  
corr1 <- model.matrix(~0+., data=df)%>%
  cor(use="pairwise.complete.obs")%>%
  ggcorrplot(show.diag = F, type = "lower", lab=TRUE, lab_size = 2)
corrMartrix <- cor(df)
corr1

corrMartrix
```


```{r}
#Now we are going to create a correlation model to identify questions that might indicate a relationship. To do this we need to standardize all response using a number. 

## Yes and affirmative responses like "lumped", "categorized", and "clearly detailed" will be equal to 1.
## No or "absent" for question 12 will be equal to 0.
## Null values will not be considered. 

# First we will start by using the data frame with only Cumulative Effects analyzed.

CEonly=datsheet%>%
  filter(`Resource_4`== "YES")

CEonly[CEonly=="YES"]='1'
CEonly[CEonly=="CATEGORIZED"]='1'
CEonly[CEonly=="LUMPED"]='1'
CEonly[CEonly=="CLEARLY DETAILED"]='1'
CEonly[CEonly=="NO"]='-1'
CEonly[CEonly=="ABSENT"]='-1'
CEonly[CEonly=="N/A"]='0'

#Jim's helpfunction no longer needed since DF changed to 1 and 0

#subset datasheet for columns 5 through 10 and change YES to 1 everything else to 0 -- Aimee's new df w/ only 1/0 values
library(dplyr)

df <- CEonly%>% 
  select('5','6','7','8','9', '10','22')%>%
  filter('5'!="N/A")%>%
  mutate_all(~as.numeric(.))

View(df)
corr1 <- model.matrix(~0+., data=df)%>%
  cor(use="pairwise.complete.obs")%>%
  ggcorrplot(show.diag = F, type = "lower", lab=TRUE, lab_size = 2)
corrMartrix <- cor(df)
corr1

```
